---
title: "Markdown Take 3"
author: "Connor burleigh"
date: "April 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
```

#Analysis of Personal Activity Data
##Introduction
The data being analyzed is a the number of footsteps taken by an individual collected at 5 minute intervals over the course of October and November, 2012. 

```{r }
activity_data<-read.csv("activity.csv")
summary(activity_data)
```

##Extracting Daily Total
The data is classified by date, using which we can extract the daily totals and summary stats and daily means and medians, and then providing a histogram.

```{r }
dates<-activity_data[,2]
daily_totals<-as.vector((by(activity_data$steps,dates,sum)))
hist(daily_totals,main="Daily Steps",xlab="Steps",col="grey",border="black",breaks = seq(from=0,to=25000,by=1000))

clean_daily_totals<-(daily_totals[!is.na(daily_totals)])
median(clean_daily_totals)
mean(clean_daily_totals)

```


##Removing NAs and Creating a Time Series
```{r }
no_nas<-activity_data[-which(is.na(activity_data$steps)),]
interval_averages<-as.vector(by(no_nas$steps,no_nas$interval,mean))

times<-as.POSIXlt(seq(ymd_hm("2000-01-01 00:00"),ymd_hm("2000-01-01 23:55"),by='5 mins'))

time_series_plot_data<-data.frame(times,interval_averages)
plot(time_series_plot_data,ylab="Average Steps",xlab="Time",main="Average Steps Per Interval",type="l")
time_series_plot_data$times<-strftime(time_series_plot_data$times,format="%H:%M")
time_series_plot_data[which.max(time_series_plot_data$interval_averages),]

```

##Replacing the NAs
```{r }
clean_activity_data<-activity_data
Interval_Average<-rep(interval_averages,61)
clean_activity_data<-cbind(clean_activity_data,Interval_Average)
clean_activity_data$steps[is.na(clean_activity_data$steps)]<-clean_activity_data$Interval_Average[is.na(clean_activity_data$steps)]

adjusted_daily_totals<-as.vector(by(clean_activity_data$steps,dates,sum))
hist(adjusted_daily_totals,main="Adjusted Daily Steps",xlab="Steps",col="grey",border="black",breaks = seq(from=0,to=25000,by=1000))

median(adjusted_daily_totals)
mean(adjusted_daily_totals)

```

##Weekdays vs Weekends
```{r }
day<-as.factor(weekdays(as.POSIXct.Date(clean_activity_data$date)))
clean_activity_data2<-cbind(clean_activity_data,day)
Day_Index<-c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")
Weekday_Index<-c("Weekday","Weekday","Weekday","Weekday","Weekday","Weekend","Weekend")
clean_activity_data2$weekday_weekend<-Weekday_Index[match(clean_activity_data2$day,Day_Index)]
weekday_weekend<-as.factor(clean_activity_data2$weekday_weekend)

weekday_data<-subset(clean_activity_data2,weekday_weekend=="Weekday")
weekday_intervals<-as.vector(by(weekday_data$steps,weekday_data$interval,mean))
weekday_plot_data<-data.frame(times,weekday_intervals)
weekend_data<-subset(clean_activity_data2,weekday_weekend=="Weekend")
weekend_intervals<-as.vector(by(weekend_data$steps,weekend_data$interval,mean))
weekend_plot_data<-data.frame(times,weekend_intervals)

par(mfrow=c(1,2))
plot(weekday_plot_data,type="l",ylab="Average Steps",xlab="Time",main = "Average Weekday Steps")
plot(weekend_plot_data,type="l",ylab="Average Steps",xlab="Time",main = "Average Weekend Steps")
```

